---
title: "Descriptive Statistics IMDB Dataset"
author: "Mahesh Gurumoorthi"
date: "2025-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
# install.packages("pastecs")
library(pastecs)
# install.packages("lattice")
library(lattice)
options(scipen = 900)


```

## Problem Statement of Descriptive Statistics using Amazon TV Shows (IMDB)
The problem statement focuses on analyzing IMDb scores by leveraging details from IMDb votes found in a Kaggle dataset. IMDb scores represent the overall user rating of a movie, typically ranging from 1 to 10, based on individual user votes. These votes capture audience reactions, preferences, and perceptions of a film’s quality.

The objective of this analysis is to examine the relationship between IMDb votes and the assigned IMDb scores, identifying key patterns, trends, and influencing factors that contribute to a movie’s final rating. This includes:

Evaluating the distribution of votes across different movies and genres.

Assessing how the number of votes impacts the IMDb score—whether higher vote counts lead to more stable ratings.

Identifying anomalies, such as movies with exceptionally high or low scores relative to their vote distribution.

Understanding possible biases in user ratings based on factors like genre, popularity, or external reviews.

Exploring statistical techniques, including correlation analysis and predictive modeling, to estimate IMDb scores from vote-related metrics.

By conducting this study, valuable insights can be gained into how IMDb ratings are shaped by user votes. Additionally, the findings may help predict movie ratings based on voting patterns, improving future recommendations and audience engagement strategies.


```{r cars}
dataset <- read.csv('/Users/maheshg/Dropbox/Sample Datasets Kaggle/Amazon Prime TV Shows/titles.csv')
clean_dataset <- na.omit(dataset)
# print(clean_dataset)
colnames(clean_dataset)
str(clean_dataset)
###Finding the values exist in the dataset? 
which(is.na(clean_dataset))
sum(is.na(clean_dataset))
anyNA(clean_dataset) ### This is cleaned dataset omitted na values 
anyNA(dataset) ### There are some NA values exists in the maindataset 

summary(clean_dataset)
str(clean_dataset)
by(clean_dataset,clean_dataset$release_year,summary)


stat.desc(clean_dataset$imdb_score)
stat.desc(clean_dataset$imdb_votes)

### Coefficient Variation : 
sd(clean_dataset$imdb_score) / mean(clean_dataset$imdb_score)
sd(clean_dataset$imdb_votes) / mean(clean_dataset$imdb_votes)

tab <- table(clean_dataset$imdb_score)
print(tab)
tab1 <- table(clean_dataset$imdb_votes)
print(tab1)
sort(table(clean_dataset$imdb_votes),decreasing = TRUE)
sort(table(clean_dataset$imdb_score),decreasing = TRUE)

summary(clean_dataset)


###Identifying the correlation between two variables imdb scores and imdb votes 
cor(clean_dataset$imdb_score,clean_dataset$imdb_votes)
###Creating contingency table 
median(clean_dataset$imdb_votes)
mean(clean_dataset$imdb_votes)
clean_dataset$imdb_votes_size <- ifelse(clean_dataset$imdb_votes < median(clean_dataset$imdb_votes),
                          "average","beyond average")
table(clean_dataset$imdb_votes_size)
clean_dataset$imdb_score_size <- ifelse(clean_dataset$imdb_score < median(clean_dataset$imdb_score),
                          "average","beyond average")
table(clean_dataset$imdb_score_size )
table(clean_dataset$imdb_votes_size,clean_dataset$imdb_score_size )
prop.table(table(clean_dataset$imdb_votes_size,clean_dataset$imdb_score_size ))
round(prop.table(table(clean_dataset$imdb_votes_size,clean_dataset$imdb_score_size ),1),2) ### Percentage by Row
round(prop.table(table(clean_dataset$imdb_votes_size,clean_dataset$imdb_score_size ),2),2) ### Percentage by Column



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

###Creation of Mosaic Plot : 
mosaicplot(table(clean_dataset$imdb_votes_size,clean_dataset$imdb_score_size ),
           color = TRUE,
           main = "Mosaic Plot View between IMDB Votes vs IMDB Scores",
           xlab = "Votes",
           ylab = "Size")

###Creation of Bar Plot : 
barplot(table(clean_dataset$imdb_votes_size))
barplot(prop.table(table(clean_dataset$imdb_votes_size)))

###Plotting the value 
ggplot(clean_dataset)+
  aes(x=imdb_score_size) +
  geom_bar()

###Histogram 
str(clean_dataset$imdb_score_size)
hist(clean_dataset$release_year,
     main = "Histogram View of all the IMDB Shows based on the release year ")
ggplot(data = clean_dataset) +
  aes(x = clean_dataset$release_year) +
  labs(title = "Plotting the release year imdb shows") +
  # annotate("This gives negative skewed view based on the distribution of the dataset") +
  geom_histogram()

str(clean_dataset)

###Remove the outliers in the release year attribute 
Q1 <- quantile(data = clean_dataset$release_year,0.25)
Q3 <- quantile(data= clean_dataset$release_year,0.75)
IQR <- Q3 - Q1

lower_bound = Q1 - 1.5 * IQR
upper_bound = Q3 + 1.5 * IQR

# clean_dataset$score_details <- clean_dataset$imdb_score[(clean_dataset$imdb_score >= lower_bound) & (clean_dataset$imdb_score <= upper_bound)]


###Boxplot details of IMDB Score and IMDB Votes:

# boxplot(clean_dataset$imdb_score ~ clean_dataset$imdb_votes)
par(mfrow = c(2,2))

boxplot(clean_dataset$imdb_score)
boxplot(clean_dataset$imdb_votes)

dotplot(clean_dataset$imdb_votes ~ clean_dataset$imdb_score)

###Stack View Plot : 
ggplot(datat = clean_dataset) + 
  aes(x = clean_dataset$imdb_votes, 
      y = clean_dataset$imdb_score) +
  geom_dotplot(binaxis = "y", stackdir = "center")

### Scatter Plot : 
par(mfrow = c(1,1))
plot(clean_dataset$imdb_score, clean_dataset$imdb_votes)

ggplot(data = clean_dataset) +
  aes(x = clean_dataset$imdb_score, 
      y = clean_dataset$imdb_votes) +
  geom_point() #### Showing strong positive association between two variables 

ggplot(clean_dataset) +
  aes(x = imdb_votes, 
      y = imdb_score,
      colour = production_countries) +
  labs(title = "Scatter Plot Details based on the IMDB Scores and IMDB Votes grouped by Production Countries") +
  geom_point() +
  scale_color_hue()


###Line Plot : 
plot(clean_dataset$imdb_score,
     type = "l")

### QQ Plot : 
qqnorm(clean_dataset$imdb_score)
qqline(clean_dataset$imdb_votes)

### QQ Plot with confidence from the car package : 
qqplot(clean_dataset$imdb_votes,clean_dataset$imdb_score,plot.it = TRUE)

qqplot(clean_dataset$imdb_score,clean_dataset$imdb_votes, groups = clean_dataset$imdb_score_size)

plot(density(clean_dataset$imdb_score))

ggplot(clean_dataset) +
  aes(x = clean_dataset$imdb_score) +
  geom_density()
 



# plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
